---
title: "Homework 1"
author: "AIDAN HOLLAND"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Instructions

**Exercises:** 1-5 (Pgs. 6-7); 1-2, 5 (Pg. 12); 1-5 (Pgs. 20-21); Open Response

**Submission:** Submit via an electronic document on Sakai. Must be submitted as a HTML file generated in RStudio. All assigned problems are chosen according to the textbook *R for Data Science*. You do not need R code to answer every question. If you answer without using R code, delete the code chunk. If the question requires R code, make sure you display R code. If the question requires a figure, make sure you display a figure. A lot of the questions can be answered in written response, but require R code and/or figures for understanding and explaining.

```{r, include=FALSE}
library(tidyverse)
library(ggplot2)
```

# Chapter 1 (Pgs. 6-7)

##  Exercise 1
```{r}
ggplot(data=mpg)
```

I see absolutely nothing. There is just a blank space for a graph. Why am I even doing this nonsense?

## Exercise 2
```{r}
dim(mpg)
nrow(mpg)
ncol(mpg)
```
There are 234 rows and 11 columns in the dataset **mpg**.

##  Exercise 3
```{r, eval=FALSE}
?mpg
```

```{r}
unique(mpg$drv)
```


The variable **drg** is a factor variable that takes the following values:

- "f" = front-wheel drive
- "r" = rear-wheel drive
- "4" = 4-wheel drive

##  Excercise 4
```{r}
ggplot(data=mpg,aes(x=hwy,y=cyl)) +
  geom_point() + 
  xlab("Highway Miles Per Gallon") +
  ylab("Number of Cylinders")
```


##  Excercise 5
```{r}
ggplot(data=mpg,aes(x=class,y=drv)) + 
  geom_point() + 
  xlab("Type of Car") +
  ylab("Type of Drive")
```

Scatter plots are not meant to visualize the relationship between two categorical/qualitative variables. 

# Chapter 1 (Pg. 12)

##  Exercise 1
What has gone wrong is the fact that the "color = 'blue'" code is within the aes() code. ggplot aes() will automatically create a scale of colors, shapes, etc. for a particular class if the code is within it, but if the desired outcome was for the points to be blue, it should be outside of aes().

##  Exercise 2
The categorical variables are the following, as they are stored as string values:
manufacturer
model
trans
drv
fl
class

The continuous variables are the following, as they are stored as integers or doubles:
displ
cyl
cty
hwy

I believe year may serve as either depending on context. When you run the command mpg, the tops of the columns will display the variables and their object type.

##  Exercise 5
```{r}
ggplot(data=mpg) +
  geom_point(mapping = aes(x=displ, y=cty), stroke = 1, color = "blue", fill = "green", shape = 24, size = 2)
```

The stroke aesthetic changes the thickness of the border given a shape that can have a color and a fill. This works with the filled shapes (21-24) as they can have both a border of a color and a fill of a color.

# Chapter 1 (Pgs. 20-21)

##  Exercise 1
Line chart - geom_smooth()
Boxplot - geom_boxpot()
Histogram - geom_histogram()
Area chart - geom_area()

##  Exercise 2
```{r}
ggplot(
 data = mpg,
 mapping = aes(x = displ, y = hwy, color = drv)
) +
 geom_point() +
 geom_smooth(se = FALSE)
```

I predict the output will be a scatterplot graph of engine displacement vs. highway mpg, colored by the class of drivetrain. The different classes will have their own line plots as well.

##  Exercise 3
```{r}
ggplot(
 data = mpg,
 mapping = aes(x = displ, y = hwy, color = drv),
) +
 geom_point(show.legend = FALSE) +
 geom_smooth(show.legend = FALSE, se = FALSE)
```

show.legend = FALSE removes the legend at right of the plot. If you remove it, it is no longer there. The argument may have been used earlier in the chapter in order to reduce clutter, or maybe the exact class of drivetrain was not important to the data visualization in that moment.

##  Exercise 4
```{r}
ggplot(
 data = mpg,
 mapping = aes(x = displ, y = hwy, color = drv),
) +
 geom_point(show.legend = FALSE) +
 geom_smooth(show.legend = FALSE)
```

The argument se = FALSE removes the line plot accounting for standard error. By removing that argument, you choose to include the standard error, which may produce clutter or information insignificant to the data visualization/analysis.

##  Exercise 5

I don't know if they will look different. Let me check.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```

They do not look different. I am incredibly surprised. They do not look different because the arguments for ggplot() can be applied to geom_point geom_smooth distributively or they can be placed inside the geom_point() and geom_smooth() functions respectively to produce the same result.

# Open Response

For this exercise, use the **diamonds** dataset in the tidyverse. Use `?diamonds` to get more information about the dataset.

## Step 1: Select 1 numeric variable and 2 categorical variables. Create a graphic using `geom_boxplot()` and `facet_wrap` to illustrate the empirical distributions of the sample.
```{r}
ggplot(data = diamonds) +
  geom_boxplot(aes(x = color, y = price)) +
  facet_wrap(cut~.) +
  xlab("Color of Diamond (D-J)") + ylab("Price ($)") +
  theme_light()+ggtitle("Diamond Value, Organized by Color, Separated by Cut Quality") +
  theme_classic()
```

## Step 2: Choose 2 numeric variables and 2 categorical variables and creatively illustrate the relationship between all the variables. 
```{r}
ggplot(diamonds) +
  geom_point(aes(x=carat, y = price, color = clarity)) +
  xlab("Carat (weight)") + ylab("Price ($)") +
  ggtitle("Diamond Value vs. Carat, Colored by Clarity, Separated by Cut") +
  facet_wrap(cut~.)
```

Some noticeable trends: it's not exactly creative to consider the reality that an increasing carat will result in an increasing price. However, IF being the best clarity, it seems that the value of the diamond does not have to depend so much on that high carat and instead can have a similar price to the heaviest diamond given that the clarity is desirable. Additionally, it seems a fair diamond will have to weigh a lot more and/or be a lot more clear to achieve the same value as a diamond of a higher cut that may be not as clear or of as great of a weight.