---
title: "Homework 3"
author: "AIDAN HOLLAND"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Instructions

**Exercises:**  1,3 (Pgs. 90-91); 1 (Pg. 93); 2,4 (Pg. 99); 1,2 (Pg. 101); 2,3,5 (Pg. 104)

**Submission:** Submit via an electronic document on Sakai. Must be submitted as a HTML file generated in RStudio. All assigned problems are chosen according to the textbook *R for Data Science*. You do not need R code to answer every question. If you answer without using R code, delete the code chunk. If the question requires R code, make sure you display R code. If the question requires a figure, make sure you display a figure. A lot of the questions can be answered in written response, but require R code and/or figures for understanding and explaining.

```{r, include=FALSE}
library(tidyverse)
```

# Chapter 5 (Pgs. 90-91)

##  Exercise 1
```{r}
summary(select(diamonds,x,y,z))
ggplot(diamonds)+
  geom_histogram(mapping=aes(x=x),binwidth=0.05)
ggplot(diamonds)+
  geom_histogram(mapping=aes(x=y),binwidth=0.05)
ggplot(diamonds)+
  geom_histogram(mapping=aes(x=z),binwidth=0.05)
```

The mean size of x and y is about the same, whereas z is slightly smaller. However, the maximum of z and y are far higher than that of x. x and y mostly follow the same summary statistics until the max. It is clear that there are many more diamonds with small dimensions than there are with large dimensions. In my head, using x and y would be a look from the top down, which makes sense as diamonds are often circular or some sort of hexagonal shape, and the z is the depth along that top down axis.

##  Exercise 3
```{r}
diamonds %>%
  filter(carat>=0.99,carat<=1) %>%
  count(carat)
```

There are many more 1 carat than 0.99 carat diamonds. This could be due to rounding, maybe to be able to say that my diamond I'm trying to sell is at least 1 carat, and not less. 

# Chapter 5 (Pg. 93)

##  Exercise 1

Missing values are removed from histograms, whereas in bar plots they receive their own grouping. This makes sense because NA can be treated as a categorical, and grouped as such in a bar plot, but as it does not have any numerical data associated with it, it doesn't make sense to include in a histogram, which bins numeric data by the values.

# Chapter 5 (Pg. 99)

##  Exercise 2
```{r}
#
```

##  Exercise 4
```{r}
#
```

# Chapter 5 (Pg. 101)

##  Exercise 1
```{r}
#
```

##  Exercise 2
```{r}
#
```

# Chapter 5 (Pg. 104)

##  Exercise 2
```{r}
#
```

##  Exercise 3
```{r}
#
```

##  Exercise 5
```{r}
#
```
