---
title: "Homework 7"
author: "AIDAN HOLLAND"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Instructions

**Exercises:**  1,4 (Pg. 358); 1,4 (Pgs. 371)

**Submission:** Submit via an electronic document on Sakai. Must be submitted as a HTML file generated in RStudio. All assigned problems are chosen according to the textbook *R for Data Science*. You do not need R code to answer every question. If you answer without using R code, delete the code chunk. If the question requires R code, make sure you display R code. If the question requires a figure, make sure you display a figure. A lot of the questions can be answered in written response, but require R code and/or figures for understanding and explaining.

```{r, include=FALSE}
library(tidyverse)
library(modelr)
```


# Chapter 18 (Pg. 358)

##  Exercise 1
```{r}
sim1loess <- loess(y ~ x, data = sim1)
sim1lm <- lm(y ~ x, data = sim1)
loess <- sim1 %>%
  add_predictions(sim1loess)
sim1 <- sim1 %>%
  add_residuals(sim1lm) %>%
  add_predictions(sim1lm) %>%
  add_residuals(sim1loess, var ="resid_loess") %>%
  add_predictions(sim1loess,var = "pred_loess")
plotloess<-ggplot(sim1,aes(x=x,y=y))+
  geom_point()+
  geom_line(aes(x=x,y=pred),data=loess)
plotloess
plotloess2<-plotloess+
  geom_smooth(method="loess",se=F)
plotloess2
```

Same result!

##  Exercise 4
```{r}
sim1_2<-lm(y~x,data=sim1)
sim1<-sim1%>%
  add_residuals(sim1_2)
ggplot(sim1,aes(x=abs(resid)))+
  geom_freqpoly()
```

Looking at the frequency polgyon of absolute residuals gives a solid idea of the variation that the residuals have, but we're entirely missing out on the sign that might indicate a possible preferred over or underestimation of the residuals.

# Chapter 18 (Pg. 371)

##  Exercise 1
```{r}
model2<-lm(y~x,data=sim2)
model2a<-lm(y~x-1,data=sim2)

check<-sim2%>%
  data_grid(x) %>%
  spread_predictions(model2,model2a)
check
```

They're identical

##  Exercise 4
```{r}
model1<-lm(y~x1+x2,data=sim4)
model2<-lm(y~x1*x2,data=sim4)
sim4mods<-gather_residuals(sim4,model1,model2)
ggplot(sim4mods,aes(x=resid,color=model))+
  geom_freqpoly()+
  geom_rug()
ggplot(sim4mods,aes(x=abs(resid),color=model))+
  geom_freqpoly()+
  geom_rug()
```
