---
title: "Lab 5: Exploratory Data Analysis"
author: "Aidan Holland"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
# This is a good place to put libraries required for using the ggplot function
knitr::opts_chunk$set(echo = TRUE)
options(scipen=999)
library(tidyverse)    #Loads the tidyverse suite of packages
library(Ecdat)        #Contains Wage Data
library(knitr)        #kable() for HTML Tables
library(rainbow)
library(rrcov)
```

# Introduction

In this lab assignment, make sure you work in order of the code chunks, and **knit** after you complete each code chunk.

Consider the dataset `Wages1` from the `Ecdat` package.  
```{r,echo=F}
head(Wages1)
```

This observational dataset records the years experienced, the years schooled, the sex, and the hourly wage for 3,294 workers in 1987. *A Guide to Modern Econometrics* by Marno Verbeek utilizes this data in a linear regression context. According to Marno Verbeek, this data is a subsample from the US National Longitudinal Study. The purpose of this tutorial is to practice the creative process in exploratory data analysis of asking questions and then investigating those questions using visuals and statistical summaries. 

As a member of the birth class of 1988, I do not have any clue of what the workforce looked like in 1987. It is your job to apply your detective skills to  the information hidden in this data. For future use, utilize the modified dataset`wage` according to the R code below:

```{r}
wage=as_tibble(Wages1) %>%
  rename(experience=exper) %>%
  arrange(school)
head(wage)
```

# Part 1: Questions About Variation

### 1.1: Use a figure to find what is the most common  number of years of experience found in the data (0.5 Points).

First, use `geom_bar()` to investigate the distribution of level of experience found in `wage`.
```{r}
ggplot(wage, aes(x = experience)) +
  geom_bar() +
  labs(x = "Years of Experience", y = "Frequency") +
  ggtitle("Experience Distribution")
```

### 1.2: Use a table to find what is the most common  number of years of experience found in the data (1.5 Points). 

Use `group_by(experience)` along with the pipe `%>%` to output the most common amount of years of experience along with the number of occurrences found in the data. The most common value for years of experience is 9 and occurs 654 times. Fill in the blanks with the correct answers, and change `eval=FALSE` to `eval=TRUE` and print out the output that led you to your answer.

```{r,eval = TRUE}
wage %>%
  group_by(experience) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  head(5)
```

### 1.3: Use a figure to find what is the maximum number for years of schooling found in the data. (0.5 Points)

First, use `geom_bar()` to visualize the overall distribution of level of schooling found in the data.

```{r}
ggplot(wage, aes(x = school)) +
  geom_bar() +
  labs(x = "Years of Schooling", y = "Frequency") +
  ggtitle("Level of Schooling Distribution Distribution")
```

### 1.4: Use a table to find what is the maximum number for years of schooling found in the data. (1.5 Points)

Next, modify the code in Question 1.2 to display the maximum level of schooling and the number of workers in the data that had that number of schooling. The maximum number of years in school was 16 years which occurred 16 times in our sample. Fill in the blanks with the correct answers.

```{r}
wage %>%
  group_by(school) %>%
  summarise(count = n()) %>%
  arrange(desc(school)) %>%
  head(5)
```


# Part 2: Questions about Covariation

### 2.1: Follow-up to Questions 1-2: Is there a relationship between level of schooling and level of experience? (1 Point)

Use `geom_point()` to display a scatter plot representing the relationship between these two discrete numeric variables. Consider using `alpha=0.1` to indicate where the relationship is represented the best.

The years of experience seem to decrease (increase/decrease) as the years of schooling increases. Is this what you expected to see? Yes (yes/no). 

```{r}
ggplot(wage, aes(x = school, y = experience)) +
  geom_point(alpha = 0.1) +
  labs(x = "Schooling (years)", y = "Experience (years)") +
  ggtitle("Experience vs. Schooling")
```

Question: Practically, what reasons do you hypothesize for this observed relationship? Write your answer in complete sentences below:

Over the course of a career, it is very possible that one could decide to trade off school for work, or work for school. If you have more years of schooling, be that undergrad, postbacc, grad, postdoc, or others, you give up years of experience that you could have acquired had you chosen to stop schooling at some point. Similarly, if you choose to finish up schooling at 6 years for example, you're likely to have had a lot more time to gain field experience.

### 2.2: How do hourly wages differ between males and females? (1 Point)

Use `geom_freqpoly()` to compare the distribution of wage of females to the distribution of wage of males.

```{r}
ggplot(wage, aes(x = wage, color = sex)) +
  geom_freqpoly() +
  labs(x = "Wage", y = "Frequency") +
  ggtitle("Wage Distribution by Gender")
```


Question: Where do these distributions look the same and/or where do they differ? Write your answer in complete sentences below:

The distributions look almost identical in shape, indicating a wage distribution that is heavily skewed right as those that make a very high wage are in low quantity. However, for every male wage frequency, it seems that the female equivalent is about 1 or 2 levels behind in wage. The highest frequency for both sexes is around 350, but it is around 4-6 for males, and 3-5 for females. The discrepancy only grows as wage increases, and many of the highest levels of wage don't have female data (e.g. the gap between ~33 and ~39) while males have data for every point.

### 2.3: How do hourly wages differ between males and females? (1.5 Points)

Use `group_by()` along with summarize to report the mean `wage`, standard error of `wage`, and 95% confidence interval for the unknown population mean hourly wage for the various levels of `sex`. The standard error is equal to the standard deviation divided by the square root of the sample size. The 95% confidence interval is approximated by obtaining the lower and upper bound of an interval within 2 standard errors of the sample mean. Based on the confidence limits, do we have statistical evidence to say that the average hourly wage for men was different than the average hourly wage for women in 1987? ______ (yes/no). 


```{r,eval=FALSE}
wage %>% 
  group_by(sex) %>%
  summarize(n=n(),mean= mean(wage),se=sd(wage)/sqrt(n),
            lb=mean-2*se,ub=mean+2*se)
```

Question: How would you explain your answer in terms of the confidence intervals that are constructed below? Write your answer in complete sentences below:

For females, the 95% confidence interval ranges from 5.001698-5.292149. For males, it is 6.144536-6.481507. The complete lack of overlap suggests that there is a significant difference in the wage gap in 1987.

```{r}
wage %>% 
  group_by(sex) %>%
  summarize(n=n(),mean=mean(wage),se=sd(wage)/sqrt(n),
            lb=mean-2*se,ub=mean+2*se)
```


### 2.4: Does the relationship between hourly wage and years of experience differ between the sexes? (0.5 Points)

Use `geom_point()` along with the option `color=sex` to overlay scatter plots. Does there seem to be a clear distinction between female and male regarding this relationship? yes (yes/no).

```{r}
ggplot(wage, aes(x = experience, y = wage)) +
  geom_point(aes(color = sex)) +
  labs(x = "Years of Experience", y = "Hourly Wage", color = "Sex") +
  ggtitle("Hourly Wage vs. YoE") +
  facet_wrap(~ sex)
```
 
### 2.5: Does the relationship between hourly wage and years of schooling differ between the sexes? (0.5 Points)

Repeat the graphic created in Question 4 replacing `x=experience` with `x=school`. Does there seem to be a clear distinction between female and male regarding this relationship? yes (yes/no).

```{r}
ggplot(wage, aes(x = school, y = wage)) +
  geom_point(aes(color = sex)) +
  labs(x = "Years of School", y = "Hourly Wage", color = "Sex") +
  ggtitle("Hourly Wage vs. Schooling") +
  facet_wrap(~ sex)
```

### 2.6: What is the relationship between hourly wage and the interaction between the years of experience and years of schooling? (0.5 Points)

The graphic below summarizes the average hourly wage for the different combinations of schooling and experience level. The additional `facet_grid(~sex)` makes comparing the relationship of the three key numeric variables between the sexes quite easy. 

```{r}
wage %>%
  group_by(experience,school,sex) %>%
  summarize(n=n(),mean=mean(wage)) %>%
  ungroup() %>%
  ggplot() +
    geom_tile(aes(x=experience,y=school,fill=mean)) +
  scale_fill_gradientn(colors=c("black","lightskyblue","white"))+
    facet_grid(~sex) + theme_dark()
```

Question: What are some differences between the sexes regarding this relationship that are apparent in this chart? Write your answer in complete sentences below:

The area that represents the bottom and left of both graphs (indicating relatively low years of experience, schooling, or both) seems to be a big difference. Females clearly have a lower mean wage in these areas, with many dark squares, while males have a higher mean wage. Additionally, the peak mean wage for males is higher than females. Males maintain a consistently high mean wage as years of school decrease and years of experience increase, whereas females seem to fall off in mean wage faster and sooner as females don't even have recorded data for unschooled, high experienced individuals.

### 2.7: What is the relationship between hourly wage and the interaction between the years of experience and years of schooling? (1 Point)

The next figure is similar to the previous one except that the tile color reflects the standard deviation of wage rather than the mean. Interactions of experience and school levels containing less than or equal to 10 instances are ignored in this image. 

```{r}
wage %>%
  group_by(experience,school,sex) %>%
  summarize(n=n(),sd=sd(wage)) %>%
  ungroup() %>%
  filter(n>10) %>%
  ggplot() +
  geom_tile(aes(x=experience,y=school,fill=sd)) +
  scale_fill_gradientn(colors=c("black","lightskyblue","white"))+
  facet_grid(~sex) + theme_dark()
```

Question: Which plot is generally darker and what does that imply? Write your answer in complete sentences below:

The female plot is generally darker, implying that there is not a lot of variability in the wage at those data points. A lighter square would imply that there is a high level of variability in the wage.

Question: Specifically for the scenario where a worker has 5 years of experience and 11 years of schooling, what does the extreme contrast between female and male cells imply for this figure? Write your answer in complete sentences below:

Female wages have a far higher tendency to cluster or possibly even be boxed inside an expected value for individuals with that level of experience and schooling. Male wages for that level have a much higher variation, which is not necessarily good or bad. However, it may indicate that potential earnings are higher for men, possibly because of gender disparities.